---
title: "2 - MWE"
output: html_document
---

```{r setup, include=FALSE}
# I like to put all of my attached libraries up here
library(tidyverse)

# Make a color palette
colorRamp <- viridisLite::viridis(n=2)

knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors=FALSE)
```

# Minimum Working Examples - MWE

In *1-Basic.Rmd* we used a default dataset (cars) included in R to test drive Rmarkdown. But let's try creating our own test data. Much of this material is taken from Jared Knowles' blog: <https://www.jaredknowles.com/journal/2013/5/27/writing-a-minimal-working-example-mwe-in-r>.

Following, we will make a basic data frame of simulated pollock lengths.

```{r data}
dataRows <- 250
lengthData <- data.frame(id = seq(1, dataRows), 
                        sex = sample(c("male", "female"), size = dataRows, replace = TRUE), 
                        length = abs(rnorm(dataRows, mean = 357, sd = 169))
)

head(lengthData)
```

# Make a table

There are several packages that help you format tables (<https://rmarkdown.rstudio.com/lesson-7.html>). The function *kable* is included in the package **knitr** which is already an Rmarkdown dependency.

```{r table1}
summary <- lengthData %>%
   group_by(sex) %>%
   summarize(Mean = mean(length),
             SD = sd(length),
             Count = n()
   ) 

knitr::kable(summary)

```

<br>

Here is an example of in-line R code: there are `r dataRows` rows of data in this table.



# GGplot

Here are a couple of ggplot outputs. Note that we use the color palette created above.

```{r boxplot}
ggplot(lengthData, aes(x=sex,y=length, fill=sex)) +
   geom_boxplot() +
   scale_fill_manual(values=colorRamp) +
   theme_bw()
```

```{r histogram}
ggplot(lengthData, aes(x=length, fill=sex)) +
   geom_histogram(bins=40, position="dodge") +
   scale_fill_manual(values=colorRamp) +
   theme_bw()
```